---
output: html_document
---

#Practical Machine Learning course

##Prediction Assignment Writeup

###Task description

This is the "classe" variable in the training set. We may use any of the other variables to predict with.

###Data preprocessing

First, we need to load our data:

```{r}
library(caret)
train <- read.csv('pml-training.csv')
test <- read.csv('pml-testing.csv')
```

Second, we don't want to use all variables to build the model. Deleting variables with near zero variance may help.

```{r}
train <- train[!nearZeroVar(train, saveMetrics = TRUE)$nzv]
```

Third, we can delete variables that may be NA:

```{r}
names = character()
for(name in colnames(train))
{
    if (sum(is.na(train[[name]])) == 0)
    {
        names = c(names, name)
    }
}
train = train[names]
```

Let's look on variables names:

```{r}
colnames(train)
```

We see that first 5 variables is not important: timestamps and row number. We can delete it:

```{r}
train = train[6:ncol(train)]
```

Train and test data has different variable's types. Let's fix it:

```{r}
d = train[1:20,]
for(i in colnames(d)[1:ncol(train)-1])
{
    for(j in 1:20)
    {
        d[[i]][j] = test[[i]][j]
    }
}
test = d
```

###Training model

We'll be use SVM (Support Vector Machine) machine learning algorithm with linear kernel implemented in e1071 R package. We need only one parameter to tweak: parameter C, that control regularization.

First, load package and define function to persorm cross-validation:

```{r}
library(e1071)
folds <- createFolds(train$classe, k = 4, returnTrain = TRUE)
cv <- function(data, folds, cost){
    accuracy = 0
    for(i in 1:length(folds))
    {
        f <- folds[[i]]
        Tr <- data[f,]
        Te <- data[-f,]
        m <- svm(classe~., data = Tr, cost = cost)
        r <- predict(m, Te)
        accuracy <- accuracy + sum(r == Te$classe) / length(Te$classe)
    }
    accuracy / length(folds)
}
```

Second, calculate optimal value for C(cost) parameter:

```{r}
bestc = 1
accuracy = 0
c = c(1, 10, 100, 1000)
for(cc in c){
    acc <- cv(train, folds, cc)
    if (acc > accuracy){
        accuracy = acc
        bestc = cc
    }
    print(paste('C = ', as.character(cc), ', accuracy = ', as.character(acc)))
}
```

We get bestc value:

```{r}
bestc
```

And now we can train the model with optimal cost:

```{r}
model <- svm(classe~., data = train, cost = bestc)
```

With accuracy:

```{r}
accuracy
```

###Predicting

Let's predict values for out test set with trained model:

```{r}
predict(model, test)
```

